---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
  keep_md: true
---
  
  
## Loading and preprocessing the data
Load the data and ignore missing values for now
```{r}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
data <- read.csv('activity.csv', header = T)
data_new <- data[complete.cases(data), ]
```
  
  
## What is mean total number of steps taken per day?
First, a histogram of the number of steps is plotted below.
```{r}
steps_per_day <- aggregate(data_new$steps, list(data_new$date),sum)
png("Figure/plot1.png")
hist(steps_per_day$x, main="Total number of steps per day", xlab="steps", ylab="frequency")
dev.off()
```

The mean and median are calculated:
```{r}
mean_steps <- mean(steps_per_day$x)
median_steps <- median(steps_per_day$x)
```

The mean of the number of steps is:
```{r}
mean_steps
```
and the median is:
```{r}
median_steps
```

## What is the average daily activity pattern?
Below a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days, is plotted.
```{r}
steps_per_int <- aggregate(data_new$steps, list(data_new$interval),mean)
png("Figure/plot2.png")
plot(steps_per_int$Group.1,steps_per_int$x,type="l", main = "Mean steps per interval", xlab="interval", ylab = "steps")
dev.off()
```

The interval with the maximum average number of steps is:
```{r}
steps_per_int[which.max(steps_per_int$x),]$Group.1
```

## Imputing missing values

The number of NA's is:
```{r}
dim(data)[1]-dim(data_new)[1]
```
To fill the NA's, the mean of that interval is used.
```{r}
data_better <- data
for (i in 1:dim(data)[1]){
  if (is.na(data$steps[i])==T){
    data_better$steps[i] = steps_per_int[steps_per_int$Group.1 == data$interval[i],]$x
  }
}
```

Below a hist of the better data set is shown.
```{r}
steps_per_day_better <- aggregate(data_better$steps, list(data_better$date),sum)
png("Figure/plot3.png")
hist(steps_per_day_better$x, main="Total number of steps per day", xlab="Date", ylab="Steps")
dev.off()
```

The mean and median are calculated:
```{r}
mean_steps_better <- mean(steps_per_day_better$x)
median_steps_better <- median(steps_per_day_better$x)
```

The mean of the number of steps is:
```{r}
mean_steps_better
```
and the median is:
```{r}
median_steps_better
```

## Are there differences in activity patterns between weekdays and weekends?
First format date column to date format and add a column with weekday or weekend
```{r}
data_better$date <- as.Date(data_better$date)
data_better$day <- weekdays(data_better$date)
for (i in 1:nrow(data_better)) {
    if (data_better[i,]$day %in% c("zaterdag","zondag")) {
        data_better[i,]$day<-"weekend"
    }
    else{
        data_better[i,]$day<-"weekday"
    }
}
stepsByDay <- aggregate(data_better$steps ~ data_better$interval + data_better$day, data_better, mean)
```

Next, make the plot
```{r}
names(stepsByDay) <- c("interval", "day", "steps")
library(lattice)
png("Figure/plot4.png")
xyplot(steps ~ interval | day, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
dev.off()
```


